---
title: "eds221-day8-comp1"
format: html
editor: visual
---

## GGplot Basics Reviewed

```{r}
library(tidyverse)
library(janitor)
library(here)
library(patchwork)
library(gghighlight)
library(paletteer)
library(gapminder)
library(ggrepel)
library(sf)
```

## Read in

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))

lizards
```

## Data Wrangling

#### Subset to contain Western Whiptails, Colorado Checkered, and Eastern Fence Lizards. Add a new column w/ total length mm -\> cm Group by species and site Return a summary table of mean total length, sd, by species and site

```{r}
lizard_summarized_table <- lizards %>% 
  filter(common_name %in% c("western whiptails", "colorado checkered", "eastern fence")) %>% 
  mutate(total_length_cm = total_length / 10) %>% 
  group_by(common_name, site) %>% 
  summarize(mean_total_length_cm = mean(total_length_cm, na.rm = TRUE), sd_total_length_cm = sd(total_length_cm, na.rm = TRUE))
```

```{r}
lizard_summarized_table
```

## Data Viz w/ the Liz

```{r}
lizard_plot <- 
  ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name))
```

```{r}
lizard_plot
```

## Facet

```{r}
lizard_facet <-
  ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_wrap(~ common_name)
```

```{r}
lizard_facet
```

```{r}
lizard_facet2 <-
  ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_wrap(~ common_name, scales = "free")
```

```{r}
lizard_facet2
# scales are not uniform, can be dangerous bc audience can incorrectly compare directy across
```

```{r}
lizard_counts <- lizards %>% 
  mutate(total_lizard_count = dplyr::count(common_name))
## does not work because mutate introduced the need for non-characters
```

```{r}
lizard_counts <- lizards %>% 
  dplyr::count(common_name)
```

```{r}
lizard_counts
```

```{r}
lizard_bar
```

```{r}
lizard_bar <- ggplot(data = lizard_counts, aes(x = common_name, y = n)) +
  geom_col(aes(fill = common_name)) 
```

```{r}
lizard_bar2 <- ggplot(data = lizard_counts, aes(y = common_name, x = n)) +
  geom_col(aes(fill = common_name)) 
```

```{r}
lizard_bar2
```

## Factoring

```{r}
lizard_bar_factored <- ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) +
  geom_col() 
```

```{r}
lizard_bar_factored
```

```{r}
lizard_bar_factored2 <- ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) +
  geom_col(aes(fill = common_name)) 
```

```{r}
lizard_bar_factored2
```

## Breaks, Color Gradients

```{r}
lizard_scale <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_y_continuous(limits = c(0, 100),
                     breaks = seq(from = 0, to = 100, by = 20),
                     expand = c(0, 0)) +
  scale_color_gradient(low = "pink", high = "orange") +
  theme_minimal()
```

```{r}
lizard_scale
```

## Hline and Vline

```{r}
gapminder_plot <- ggplot(data = gapminder, aes(x = lifeExp, y = gdpPercap)) +
  geom_point(aes(color = gdpPercap)) +
  annotate("text", x = 40, y = 9000, label = "A LABEL!") +
  geom_hline(yintercept = 7000, linetype = "dotted", color = "purple") +
  geom_vline(xintercept = mean(gapminder$lifeExp))
```

```{r}
gapminder_plot
```

## Text Repel and Highlighting

```{r}
gapminder_2002 <- gapminder %>% 
  filter(year == 2002, continent == "Europe") %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(shape = 23, size = 4, color = "hotpink", fill = "plum") +
  geom_text_repel(aes(label = country)) +
  gghighlight(country == "Poland", label_key = country)

gapminder_2002
```

```{r}
gapminder_2005 <- gapminder %>% 
  filter(year == 2007) %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(shape = 23, size = 4, color = "hotpink", fill = "plum") +
  geom_text_repel(aes(label = country)) +
  gghighlight(continent == "Europe", label_key = country, max_highlight = 135)

gapminder_2005
```

## Making Maps + Paletteer 

```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(name = Name) %>% 
  ggplot() +
  geom_sf(aes(fill = name), color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
   theme(panel.background = element_rect(color = "purple", size = 5, fill = "pink"), plot.background = element_rect(fill = "yellow"))

jornada_veg
```

```{r}
plot2 <- ggplot(data = jornada_veg) +
  geom_sf(aes(fill = name), color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  theme(panel.background = element_rect(color = "purple", size = 5, fill = "pink"), plot.background = element_rect(fill = "yellow"),
        panel.grid.major = element_line(color = "magenta"),
        axis.text.x = element_text(color = "blue"))
```

```{r}
jornada_veg2 <-
  ggplot(data = jornada_veg) +
  geom_sf(aes(fill = name), color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::springfield_simpsons")

jornada_veg2
```
